---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(themis)
library(tidymodels)
library(moderndive)
library(skimr)
library(mice)
library(lubridate)
library(grid)
library(gridExtra)
```
```{r}
# Load original data
All_join <- read_csv("../KSB_621/Data/All_join.csv")

# select predetermined variables
model_df_v1 <- All_join %>%
  select(AcctNum, EventID, LineItemPriceEach, InvoiceAmount, PaymentAmount,EventDesc,
         EventTitle, EventStartDate, EventStartMonth, EventStartDayofWeek, EventTime,
         EventHours, StopHour, HoursWeekend, RoomDescription, AreaDesc, EventWOCosts, EventWOHours, RoomCapacity, ChargeClass)

# creating Event and Room categories
model_df_v2 <- model_df_v1 %>%
  filter(!grepl("Labor|Personnel Fee|Personal|Custodial Only|Custodian|Technichian|Manager|Credit|Officer|Security
              |OMIT|Meals|Beverage|Food|Lock|Key|Set up|Clean|Catering|Supervisor|Lifeguard|Application|Refund|Waived
              |Admin|Administration|Support|Maintenance", # filter data to remove non-space related data
                                 ChargeClass, ignore.case = TRUE)) %>%
  mutate(EventTitle = str_trim(str_replace_all(EventTitle, "-", " ")), # remove unwanted characters in some variables
       Event_Shift = ifelse(StopHour <= 12, "Early",
                       ifelse(StopHour > 12 & StopHour <= 17, "Midday",
                       ifelse(StopHour > 17, "Late", StopHour))), # create new feature Event Shift to identify start time frame of day
       Event_Category = ifelse(is.na(EventTitle), EventTitle,
                               case_when(
                                 str_detect(EventTitle, regex(`Religious/Congregation`, ignore_case = T)) ~ "Religious/Congregation",
                                 str_detect(EventDesc, regex(`Religious/Congregation`, ignore_case = T)) ~ "Religious/Congregation",
                                 str_detect(EventTitle, regex(`Vocational/Mentoring`, ignore_case = T)) ~ "Vocational/Mentoring",
                                 str_detect(EventDesc, regex(`Vocational/Mentoring`, ignore_case = T)) ~ "Vocational/Mentoring",
                                 str_detect(EventTitle, regex(`Social/Cultural Events`, ignore_case = T)) ~ "Social/Cultural Events",
                                 str_detect(EventDesc, regex(`Social/Cultural Events`, ignore_case = T)) ~ "Social/Cultural Events",
                                 str_detect(EventTitle, regex(`Academic/Professional`, ignore_case = T)) ~ "Academic/Professional",
                                 str_detect(EventDesc, regex(`Academic/Professional`, ignore_case = T)) ~ "Academic/Professional",
                                 str_detect(EventTitle, regex(`Sports/Fitness`, ignore_case = T)) ~ "Sports/Fitness",
                                 str_detect(EventDesc, regex(`Sports/Fitness`, ignore_case = T)) ~ "Sports/Fitness",
                                 str_detect(EventTitle, regex(`Performing/Arts`, ignore_case = T)) ~ "Performing/Arts",
                                 str_detect(EventDesc, regex(`Performing/Arts`, ignore_case = T)) ~ "Performing/Arts",
                                 TRUE ~ "Other")), # Create Event Category variable from event description and event title variables - - see key word list scripts
       Room_Category = ifelse(is.na(RoomDescription), RoomDescription,
                              case_when(
                                 str_detect(RoomDescription, regex(`Dinning/Cafeteria`, ignore_case = T)) ~ "Dinning/Cafeteria",
                                 str_detect(AreaDesc, regex(`Dinning/Cafeteria`, ignore_case = T)) ~ "Dinning/Cafeteria",
                                 str_detect(RoomDescription, regex(`Gymnasium/Gyms`, ignore_case = T)) ~ "Gymnasium/Gyms",
                                 str_detect(AreaDesc, regex(`Gymnasium/Gyms`, ignore_case = T)) ~ "Gymnasium/Gyms",
                                 str_detect(RoomDescription, regex(`Auditorium`, ignore_case = T)) ~ "Auditorium",
                                 str_detect(AreaDesc, regex(`Auditorium`, ignore_case = T)) ~ "Auditorium",
                                 str_detect(RoomDescription, regex(`Multipurpose Rooms`, ignore_case = T)) ~ "Multipurpose Rooms",
                                 str_detect(AreaDesc, regex(`Multipurpose Rooms`, ignore_case = T)) ~ "Multipurpose Rooms",
                                 str_detect(RoomDescription, regex(`Conference Rooms`, ignore_case = T)) ~ "Conference Rooms",
                                 str_detect(AreaDesc, regex(`Conference Rooms`, ignore_case = T)) ~ "Conference Rooms",
                                 str_detect(RoomDescription, regex(`Class/Lecture Rooms`, ignore_case = T)) ~ "Class/Lecture Rooms",
                                 str_detect(AreaDesc, regex(`Class/Lecture Rooms`, ignore_case = T)) ~ "Class/Lecture Rooms",
                                 str_detect(RoomDescription, regex(`General Meeting Rooms`, ignore_case = T)) ~ "General Meeting Rooms",
                                 str_detect(AreaDesc, regex(`General Meeting Rooms`, ignore_case = T)) ~ "General Meeting Rooms",
                                 str_detect(RoomDescription, regex(`Storage/Other Rooms`, ignore_case = T)) ~ "Storage/Other Rooms",
                                 str_detect(AreaDesc, regex(`Storage/Other Rooms`, ignore_case = T)) ~ "Storage/Other Rooms",
                                 str_detect(RoomDescription, regex(`Social/Cultural Events`, ignore_case = T)) ~ "Social/Cultural Events",
                                 str_detect(AreaDesc, regex(`Social/Cultural Events`, ignore_case = T)) ~ "Social/Cultural Events",
                                 str_detect(RoomDescription, regex(`Media & Computer Labs`, ignore_case = T)) ~ "Media & Computer Labs",
                                 str_detect(AreaDesc, regex(`Media & Computer Labs`, ignore_case = T)) ~ "Media & Computer Labs",
                                 str_detect(RoomDescription, regex(`Library`, ignore_case = T)) ~ "Library",
                                 str_detect(AreaDesc, regex(`Library`, ignore_case = T)) ~ "Library",
                                 str_detect(RoomDescription, regex(`Performing Arts`, ignore_case = T)) ~ "Performing Arts",
                                 str_detect(AreaDesc, regex(`Performing Arts`, ignore_case = T)) ~ "Performing Arts",
                                 str_detect(RoomDescription, regex(`Open Areas`, ignore_case = T)) ~ "Open Areas",
                                 str_detect(AreaDesc, regex(`Open Areas`, ignore_case = T)) ~ "Open Areas",
                                 str_detect(RoomDescription, regex(`Portables/Restrooms`, ignore_case = T)) ~ "Portables/Restrooms",
                                 str_detect(AreaDesc, regex(`Portables/Restrooms`, ignore_case = T)) ~ "Portables/Restrooms",
                                 str_detect(RoomDescription, regex(`Grounds and Parking`, ignore_case = T)) ~ "Grounds and Parking",
                                 str_detect(AreaDesc, regex(`Grounds and Parking`, ignore_case = T)) ~ "Grounds and Parking",
                                 str_detect(RoomDescription, regex(`Swimming Pools/Lanes`, ignore_case = T)) ~ "Swimming Pools/Lanes",
                                 str_detect(AreaDesc, regex(`Swimming Pools/Lanes`, ignore_case = T)) ~ "Swimming Pools/Lanes",
                                 str_detect(RoomDescription, regex(`Sports/Track & Field`, ignore_case = T)) ~ "Sports/Track & Field",
                                 str_detect(AreaDesc, regex(`Sports/Track & Field`, ignore_case = T)) ~ "Sports/Track & Field",
                                 TRUE ~ "Other"))) # Create Room Category variable from Room and Area Description - see key word list scripts

# Additional preprocessing
model_df_v3 <- model_df_v2%>%
   mutate(across(where(is.character) & !c(HoursWeekend), as.factor), # convert chr vars to factors
         across(where(is.Date) & !c(EventStartDate), month), # extract the month value in dates except event start date
         StartDay = day(EventStartDate), # Extract day number from event start date
         StartYr = year(EventStartDate), # Extract year from event start date
         HoursWeekend = str_replace(HoursWeekend, ":", "."),
         HoursWeekend = as.double(HoursWeekend)) %>%
  select(-AcctNum, -EventID, -EventDesc, -EventTitle, -EventStartDate, -EventTime, -ChargeClass, -RoomDescription, -AreaDesc) # remove unwanted variables


# Additional variables we decided to remove
model_df_v3.1 <- model_df_v3 %>%
  select(-LineItemPriceEach, -InvoiceAmount, -PaymentAmount, -EventWOHours, -EventHours, -StopHour)

# Summary of data
model_df_v3.1 %>%
  skim()

```

Looking at the summary statistics, there are a few issue for further considerations
- ordering of some of the factor variables - start day of the week and start day of the month
- looking at the min and max of some of the numerical, we see a number of extreme values (outliers). I will consider
  correcting for this.
- consider checking for correlation  
- Review EventTime variable. There are 900+ unique event time categories. How important is the event time start and end     
  variable to the model? I have left it out of this model for simplicity.

We have been able to review out new features creating both event and room categories, next we will attempt to see if we can address missing data in room capacity as we believe that room capacity is a key variable in answering our key business question. We have attempted imputation using predictive mean matching and are in the process of validating the predictions. 

```{r}
# Visualize missingness of the data
md.pattern(model_df_v3.1, rotate.names = T)

```


```{r}
# Attempt imputation for room capacity using adhoc imputation method.

capacity_fit <- with(model_df_v3, lm(RoomCapacity ~ LineItemPriceEach))
summary(capacity_fit)

# Attempt imputation for dataset with regression imputation
model_df_imp <- mice(model_df_v3, method = "norm.predict", m = 1, maxit = 1)
complete(model_df_imp)
```
This code imputes the missing values in the data set by the regression imputation method. The argument method = "norm.predict" first fits a regression model for each observed value, based on the corresponding values in other variables and then imputes the missing values with the predicted values.



```{r}
# Attempt at imputation for missing values using "pmm" and "logreg"   
set.seed(123)
model_df_v4  <- model_df_v3   %>%
  mice(m = 10, seed = 123) %>%
  complete(); summary(model_df_v4) 

# writing file to csv to inspect or share
model_df_v4 %>%
  write_csv("../KSB_621/Data/model_df_v4.2.rds")

model_df_v4.1 <- model_df_v4 %>%
  select(-LineItemPriceEach, -InvoiceAmount, -PaymentAmount, -EventWOHours, -EventHours, -StopHour)

model_df_v4.1 %>%
  skim()
```


```{r}
#Plotting the categories

model_df_v4 %>%
  ggplot(aes(RoomCapacity)) +
  geom_histogram() +
  theme_bw() +
  labs(title = "Room Capacity - All Time",
       x = "Room Capacity",
       y = "Count")

model_df_v4 %>%
 #count(Event_Category, sort = TRUE) %>%
  mutate(Event_Category = fct_reorder(Event_Category, RoomCapacity)) %>%
  ggplot(aes(RoomCapacity, Event_Category)) +
  geom_boxplot() +
  geom_point(aes(x = RoomCapacity, color = Event_Category), size = 0.5, show.legend = F) +
  scale_x_log10() +
  theme_bw() + 
  labs(title = "All Time",
       x = "Room Capacity",
       y = "Event Category") 

```

```{r}
model_df_v4 <- read_csv("../Data/model_df_full.rds") %>%
  mutate(across(where(is.character) & !c(HoursWeekend), as.factor), # convert chr vars to factors
         across(where(is.Date) & !c(EventStartDate), month), # extract the month value in dates except event start date
         StartDay = day(EventStartDate), # Extract day number from event start date
         StartYr = year(EventStartDate), # Extract year from event start date
         HoursWeekend = str_replace(HoursWeekend, ":", "."),
         HoursWeekend = as.double(HoursWeekend),
         EventStartDate <- NULL,
         EventTime <- NULL); model_df_v4


model_df19 <- read_csv("../Data/model_df19.rds") %>%
  mutate(across(where(is.character) & !c(HoursWeekend), as.factor), # convert chr vars to factors
         across(where(is.Date) & !c(EventStartDate), month), # extract the month value in dates except event start date
         StartDay = day(EventStartDate), # Extract day number from event start date
         #StartYr = year(EventStartDate), # Extract year from event start date
         HoursWeekend = str_replace(HoursWeekend, ":", "."),
         HoursWeekend = as.double(HoursWeekend),
         EventStartDate <- NULL,
         EventTime <- NULL); model_df19
```
LineItemPriceEach, EventStartMonth, EventStartDayofWeek, StartDay, EventHours, StopHour, EventWOCosts, EventWOHours, RoomCapacity, Event_Category, Room_Category



```{r}
#Plots distribution of room capacity

all_sample <- model_df_v4 %>%
  filter(!is.na(RoomCapacity)) %>%
  select(Event_Category, Room_Category, RoomCapacity)

all_distr_plot <- all_sample %>%
  ggplot(aes(RoomCapacity)) +
  geom_histogram() +
  theme_bw() +
  labs(title = "Room Capacity - All Time",
       x = "Room Capacity",
       y = "Count")

sample_df <- model_df19_v2.1 %>%
  filter(!is.na(RoomCapacity)) %>%
  select(Event_Category, Room_Category, RoomCapacity)

samp_distr_plot <- sample_df %>%
  ggplot(aes(RoomCapacity)) +
  geom_histogram() +
  theme_bw() +
  labs(title = "Room Capacity - 2019",
       x = "Room Capacity",
       y = "Count")

gridExtra::grid.arrange(all_distr_plot, samp_distr_plot, ncol=2)
```

Looking at the distribution of room capacity in the 2019 data, we see that the majority of room fall under the 500 room capacity bracket.


```{r}
#Plotting Event Categories

all_evnt_cat <- model_df_v4 %>%
  count(Event_Category, sort = TRUE) %>%
  mutate(Event_Category = fct_reorder(Event_Category, n)) %>%
  ggplot(aes(n, Event_Category)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_bw() + 
  labs(title = "All Time",
       x = "Count",
       y = "Event Category")


samp_evnt_cat <- model_df19_v2.1 %>%
  count(Event_Category, sort = TRUE) %>%
  mutate(Event_Category = fct_reorder(Event_Category, n)) %>%
  ggplot(aes(n, Event_Category)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_bw() + 
  labs(title = "2019",
       x = "Count",
       y = "")

gridExtra::grid.arrange(all_evnt_cat, samp_evnt_cat, ncol=2,
                        top = textGrob("All Time vs 2019 Event Category Distributions"))
```
```{r}
#Plotting Room Categories

all_room_cat <- model_df_v4 %>%
  count(Room_Category, sort = TRUE) %>%
  mutate(Room_Category = fct_reorder(Room_Category, n)) %>%
  ggplot(aes(n, Room_Category)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_bw() + 
  labs(title = "All Time",
       x = "Count",
       y = "Room_Category")


samp_room_cat <- model_df19_v2.1 %>%
  count(Room_Category, sort = TRUE) %>%
  mutate(Room_Category = fct_reorder(Room_Category, n)) %>%
  ggplot(aes(n, Room_Category)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_bw() + 
  labs(title = "2019",
       x = "Count",
       y = "")

gridExtra::grid.arrange(all_room_cat, samp_room_cat, ncol=2,
                        top = textGrob("All Time vs 2019 Room Category Distributions"))
```

```{r}
#Plotting Room Capacity by Event Category

all_plot_ev <- model_df_v4 %>%
  #count(Event_Category, sort = TRUE) %>%
  mutate(Event_Category = fct_reorder(Event_Category, RoomCapacity)) %>%
  ggplot(aes(RoomCapacity, Event_Category)) +
  geom_boxplot() +
  geom_point(aes(x = RoomCapacity, color = Event_Category), size = 0.5, show.legend = F) +
  scale_x_log10() +
  theme_bw() + 
  labs(title = "All Time",
       x = "Room Capacity",
       y = "Event Category") 

 samp_plot_ev <- model_df19_v2.1 %>%
  #count(Event_Category, sort = TRUE) %>%
  mutate(Event_Category = fct_reorder(Event_Category, RoomCapacity)) %>%
  ggplot(aes(RoomCapacity, Event_Category)) +
  geom_boxplot() +
  geom_point(aes(x = RoomCapacity, color = Event_Category), size = 0.5, show.legend = F) +
  scale_x_log10() +
  theme_bw() + 
  labs(title = "2019",
       x = "Room Capacity",
       y = "")
 



gridExtra::grid.arrange(all_plot_ev, samp_plot_ev, ncol= 2,
                        top = textGrob("All Time vs 2019 Room Capacity Distributions\n by Event Category AFTER Imputation"))
```


```{r}
#Plotting Room Capacity by Room Category

all_plot_rv <- model_df_v4 %>%
  #count(Event_Category, sort = TRUE) %>%
  mutate(Room_Category = fct_reorder(Room_Category, RoomCapacity)) %>%
  ggplot(aes(RoomCapacity, Room_Category)) +
  geom_boxplot() +
  geom_point(aes(x = RoomCapacity, color = Room_Category), size = 0.5, show.legend = F) +
  scale_x_log10() +
  theme_bw() + 
  labs(title = "All Time",
       x = "Room Capacity",
       y = "Room Category")

samp_plot_rv <- model_df19_v2.1 %>%
  #count(Event_Category, sort = TRUE) %>%
  mutate(Room_Category = fct_reorder(Room_Category, RoomCapacity)) %>%
  ggplot(aes(RoomCapacity, Room_Category)) +
  geom_boxplot() +
  geom_point(aes(x = RoomCapacity, color = Room_Category), size = 0.5, show.legend = F) +
  scale_x_log10() +
  theme_bw() + 
  labs(title = "2019",
       x = "Room Capacity",
       y = "")

gridExtra::grid.arrange(all_plot_rv, samp_plot_rv, ncol=2,
                        top = textGrob("All Time vs 2019 Room Capacity Distributions\n by Room Category AFTER Imputation"))
```
Looking at the relationship between Room Capacity and Event Category, we see that the 4000 room capacity outliers occur with Sporting, Socio-cultural, Academic or Professional events.


```{r}
#Assessing the relationship between Room Capacity and Price
cap_price <- model_df19_v2.1 %>%
  filter(!is.na(RoomCapacity),
         LineItemPriceEach > 0 & LineItemPriceEach < 2500) 
#Plot
cap_price %>%
  ggplot(aes(RoomCapacity, LineItemPriceEach)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10()
```
Checking for a relationship between Line Item Price and Room Capacity. We see a slightly positive relationship between price and capacity and would like to quantify this relationship by fitting a linear regression that attempts to predict Room capacity using price.

```{r}
# Fit regression model:
model_cap_price <- lm(RoomCapacity ~ LineItemPriceEach, data = cap_price)
# Get regression table:
get_regression_table(model_cap_price)
```
Here b0 = 258.17 (fitted intercept), b1 = 0.0213 (fitted slope for price)
So for every increase in one unit($) in price, there is an associated increase , on average, of 0.213 units(indv) of room capacity.

Question - is room capacity measured in head count or sq ft?

Test statistic -
Taking a look at the test statistic as  point estimate in relation to the hypothesis that
   H0: B1 = 0 - There is no true relationship between price and room capacity
vs HA: B1 != 0 - There is a true relationship (positive or negative) between price and room capacity

p values -
A p-value is the probability of obtaining a test statistic just as extreme or more extreme than the observed test statistic assuming the null hypothesis H0 is true.

With the p-value of our fitted regression model as zero, we can reject the null hypothesis that there is no r/ship between price and room capacity as evidence suggests that there is a significant positive relationship.

Confidence Intervals -
We’re 95% ‘confident’ that the true population slope β1 lies between 0.164 and 0.262 in addition to the associated 0.213 increase in room capacity, with a one unit increase on price. 

Conditions for using regression for inference
 -Linearity of relationship between variables - the relationship does appear linear
 -Independence of the residuals - because diff obs reference the same account, there is likely dependence between obs as a result of                                   the repeated measures for the same accounts
 -Normality of the residuals - see plot below
 
```{r}
#Testing for conditions
reg_points <- get_regression_points(model_cap_price)

reg_points %>%
  ggplot(aes(x = residual)) +
  geom_histogram(binwidth = 100, color ="steelblue") +
  #scale_x_log10() +
  labs(x = "Residual")
```
To test  for normality, we plotted the residuals and the histogram indicates that we have more negative residuals than positive. It seems our regression model's fitted Room Capacity tends to overestimate the true Room Capacity. This is further confirmed by the slight right skew.

-Equality of variance of the residuals
```{r}
reg_points %>%
  ggplot(aes(x = LineItemPriceEach, y = residual)) +
  geom_jitter(size = 0.5) +
  #scale_x_log10() +
  labs(x = "LineItemPriceEach", y = "Residual") +
  geom_hline(yintercept = 0, col = "steelblue", size = 1)
```
To test for equality of variance, we plotted the residuals against the line price gauging for dependence. There appears to be slightly more variation in the residuals for larger values of line item price but the variation is not drastic enough to suggest heteroskedasticity

```{r}

observed_slope <- cap_price %>% 
  specify(RoomCapacity ~ LineItemPriceEach) %>% 
  calculate(stat = "slope")
observed_slope


null_distn_slope <- cap_price %>% 
  specify(RoomCapacity ~ LineItemPriceEach) %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "slope")

null_distn_slope %>% 
  get_p_value(obs_stat = observed_slope, direction = "both")

visualize(null_distn_slope) +
  shade_p_value(obs_stat = observed_slope, direction = "both")
```
  
  
```{r}
# Splitting data into training and test sets
set.seed(2020)
splits <- initial_split(model_df_v5, prop = 0.6); splits

training_df <- splits %>%
  training()

testing_df <- splits %>%
  testing()
```

```{r}
model_lr <- model_df_v5 %>%
  filter(!is.na(RoomCapacity))
# Fit linear regression model

model_lr <- lm(RoomCapacity ~ ., data = model_lr)
# names(model_lr)

# Get regression table
reg_table <- get_regression_table(model_lr); reg_table

```


Considerations:
Factor variables -
1.    For event month, the default baseline is April (alphabetically),so the intercept represents the month of April. Therefore       the estimate for the other months represent the offset in relation to the month of April. 
        - is there an ideal baseline month?
2.    For event day, Friday is the baseline and represents the intercept.
3.    For event category, Academic/Professional is the baseline and represent the intercept
4.    Baselines for day and month will change when the variables are ordered.
5.    Consider interaction models based on the key continuous variables and how they interact with the factor variables

```{r}
# Checking for variable importance based on p-values assuming a level of significance of 0.05
vip::vip(model_lr, geom = c("point"))

```

```{r}
#Generalize the modelon test data
pred_model_lr <- predict(model_lr, testing_df) %>%
  as_tibble() %>%
  round(digits = 0)

rm_cap <- testing_df$RoomCapacity %>%
  as_tibble()

bind_cols(pred_model_lr, rm_cap) %>%
  filter(value...1 == value...2)

```








